---
title: "Cleaning_all_original_datasets"
author: "Madelaine, Maulid"
output:html_document: default
params:
  script_name: codes/Cleaning_all_original_datasets.Rmd
  vitek_ast_data: data/MALDITOF_RESULTS.xlsx
  biochemical_data: data/DRY_RAINY_MWZ_KILIMANJARO.xlsx
  whonet: data/WHONET DATA_9_9_2025.xlsx
  selection_file: data/DRY_RAINY_MWZ_KILIMANJARO.xlsx
  selection_sheet: DRY_RAINY_MWZ_KILIMANJARO_AST2
  kobo: "data/KOBOTOOLBOX_INIKA_SAMPLING_ANALYSIS_HUMAN_-_2025-06-12-18-24-20.xlsx"
---

```{r}
params
```

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(tidyverse)
library(readxl)
library(kableExtra)
library(here)

# Anchor the project root
here::i_am(params$script_name)
```

```{r}
#| echo: true
#| message: false
#| warning: false

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(tidyverse)
library(readxl)
library(here)

```

```{r}
## Importing the original lab data
DRY_RAINY_MWZ_KILIMANJARO <- read_excel("data/DRY_RAINY_MWZ_KILIMANJARO.xlsx",
                                        na = c("", "NA")
                                        )
# glimpse(DRY_RAINY_MWZ_KILIMANJARO)
# str(DRY_RAINY_MWZ_KILIMANJARO)
# names(DRY_RAINY_MWZ_KILIMANJARO)

```

```{r}
## Part1: Data cleaning
# Selecting relevant columns
Cleaned_Labdata_Original <- 
  DRY_RAINY_MWZ_KILIMANJARO %>%
  select(-`S/No`, -`SAMPLE TYPE`,
         -`Season Included in Number? ( evt. how?)`,-Comment_)%>%
## Replacing the NIKA_ with INIKA_ and INIKA_ to " " in INIKA_ID column
  mutate(INIKA_ID = str_replace_all(INIKA_ID, "(NIKA_)|(INIKA_)", ""))

```

Some Freezing_IDs have been misspelled

Don't think this applies anymore after the cleaning process; We detected when trying to join the labdata with whonetdata that it has been wrong written in the Freezing_ID 3 isolates 11321_1_R, 22362_2_R

We detected that there are spelling mistakes in some of the Isolates_Ids as they are not corresponding to the correct INIKA_IDs. Solution: We make a new Isolate_ID calling it Isolate_ID_Corrected, but need to keep the wrong one as well in order to manage to join the data from TVLA (VITEK)

```{r}
Cleaned_Labdata_Original<-Cleaned_Labdata_Original%>%
  mutate(
    # Extract first part of INIKA_ID
    INIKA_prefix = str_extract(INIKA_ID, "^[^_]+"),
    
    # Remove first part of Isolate_ID
    Isolate_suffix = str_replace(Freezing_ID_No, "^[^_]+_", ""),
    
    # Combine to make corrected ID
    Isolate_ID = paste0(INIKA_prefix, "_", Isolate_suffix)
  )

## Renaming the columns
# Cleaned_Labdata_Original <- 
# Cleaned_Labdata_Original %>%
 #rename(Isolate_ID = Isolate_ID_Corrected) 

# Freezing_IDS still not corrected, needs to be kept, as TVLA has used them!
Cleaned_Labdata_Original %>%
  mutate_if(is.character, factor) #%>%
  #lapply(levels)

## Re labeling the data in the columns
# List the columns to be modified

cols_to_relabel <- c(
  "COLONY MORPHOLOGY ON C3GR", 
  "COLONY MORPHOLOGY ON CARBA", 
  "COLONY MORPHOLOGY ON XLD", 
  "COLONY MORPHOLOGY ON BGA"
) 

# Apply the replacement to the selected columns
Cleaned_Labdata_Original <- 
  Cleaned_Labdata_Original %>%
  mutate(
    across(
      .cols = all_of(cols_to_relabel),
      .fns = ~ replace(., . == "NBG", "No growth")
    )
  )

## Check the homogeneity of the data registering ----
# Trick to select all the columns of type character in Cleaned_Labdata_Original dataset
Cleaned_Labdata_Original %>%
  select(where(is.character)) %>%
  # I modify all the columns from character to factor - because I want to see the "levels" : categories
  # to identify if you wrote homogeneously 
  mutate_all(factor)# %>%
  # Select one column and look at the levels
  #lapply(levels)

## Further cleaning for homogeneity entries
Cleaned_Labdata_Original <- 
  Cleaned_Labdata_Original %>%
  # `SAMPLE TYPE` is only Stool BUT I do not want to remove because I want to combine with other dataset. 
  # `SAMPLE FROM` : Fix syntax  "Adult outPatient" "Adult outpatient"
  mutate(`SAMPLE FROM` = str_replace(`SAMPLE FROM`, "Adult outPatient", "Adult outpatient")) %>%
  # `NAME OF SCHOOL/HF` Homeogeneize :  "Pasua HC"  "PasuaHC"   From To 
  mutate(`NAME OF SCHOOL/HF` = str_replace(`NAME OF SCHOOL/HF`, "PasuaHC", "Pasua HC")) %>%
  #  `COLONY MORPHOLOGY ON C3GR` "Pinkish/Reddish" <- keep this one "pinkish/Reddish"
  mutate(`COLONY MORPHOLOGY ON C3GR` = str_replace(`COLONY MORPHOLOGY ON C3GR`, "pinkish/Reddish", "Pinkish/Reddish")) %>%
  # `COLONY MORPHOLOGY ON XLD`  "Black colonies"       "Black colony"         "Blackish colonies"  need to be identical
  ## "No black colony"      "No blackish colonies" must be identical - Fixed with several choices of equality
  mutate(`COLONY MORPHOLOGY ON XLD` = 
           case_when(
             `COLONY MORPHOLOGY ON XLD`=="No blackish colonies" ~ "No black colony",
             `COLONY MORPHOLOGY ON XLD`=="Black colonies" ~ "Black colony",
             `COLONY MORPHOLOGY ON XLD`=="Blackish colonies" ~ "Black colony",
             TRUE ~ `COLONY MORPHOLOGY ON XLD`)
  ) 
```

```{r}
#Cleaned_Labdata_Original<-Cleaned_Labdata_Original%>%
 # mutate(
    # Extract first part of INIKA_ID
   # INIKA_prefix = str_extract(INIKA_ID, "^[^_]+"),
    
    # Remove first part of Isolate_ID
   # Isolate_suffix = str_replace(Freezing_ID_No, "^[^_]+_", ""),
    
    # Combine to make corrected ID
   # Isolate_ID = paste0(INIKA_prefix, "_", Isolate_suffix)
 # )

## Renaming the columns
# Cleaned_Labdata_Original <- 
# Cleaned_Labdata_Original %>%
# rename(Isolate_ID = Isolate_ID_Corrected) 

```

```{r}
# Freezing_IDS still not corrected, needs to be kept, as TVLA has used them
#Cleaned_Labdata_Original %>%
#  mutate_if(is.character, factor)# %>%
  #lapply(levels)
```

```{r}
## Re labeling the data in the columns
# List the columns to be modified

#cols_to_relabel <- c(
#  "COLONY MORPHOLOGY ON C3GR", 
#  "COLONY MORPHOLOGY ON CARBA", 
#  "COLONY MORPHOLOGY ON XLD", 
#  "COLONY MORPHOLOGY ON BGA"
#) 

# Apply the replacement to the selected columns
#Cleaned_Labdata_Original <- 
#  Cleaned_Labdata_Original %>%
#  mutate(
#    across(
#      .cols = all_of(cols_to_relabel),
#      .fns = ~ replace(., . == "NBG", "No growth")
#    )
#  )

## Check the homogeneity of the data registering ----
# Trick to select all the columns of type character in Cleaned_Labdata_Original dataset
Cleaned_Labdata_Original %>%
  select(where(is.character)) %>%
  # I modify all the columns from character to factor - because I want to see the "levels" : categories
  # to identify if you wrote homogeneously 
  mutate_all(factor)# %>%
  # Select one column and look at the levels
  #lapply(levels)

## Further cleaning for homogeneity entries
 
  Cleaned_Labdata_Original %>%
  # `SAMPLE TYPE` is only Stool BUT I do not want to remove because I want to combine with other dataset. 
  # `SAMPLE FROM` : Fix syntax  "Adult outPatient" "Adult outpatient"
  mutate(`SAMPLE FROM` = str_replace(`SAMPLE FROM`, "Adult outPatient", "Adult outpatient")) %>%
  # `NAME OF SCHOOL/HF` Homeogeneize :  "Pasua HC"  "PasuaHC"   From To 
  mutate(`NAME OF SCHOOL/HF` = str_replace(`NAME OF SCHOOL/HF`, "PasuaHC", "Pasua HC")) %>%
  #  `COLONY MORPHOLOGY ON C3GR` "Pinkish/Reddish" <- keep this one "pinkish/Reddish"
  mutate(`COLONY MORPHOLOGY ON C3GR` = str_replace(`COLONY MORPHOLOGY ON C3GR`, "pinkish/Reddish", "Pinkish/Reddish")) %>%
  # `COLONY MORPHOLOGY ON XLD`  "Black colonies"       "Black colony"         "Blackish colonies"  need to be identical
  ## "No black colony"      "No blackish colonies" must be identical - Fixed with several choices of equality
 # mutate(`COLONY MORPHOLOGY ON XLD` = 
 #          case_when(
  #           `COLONY MORPHOLOGY ON XLD`=="No blackish colonies" ~ "No black colony"#,
   #          `COLONY MORPHOLOGY ON XLD`=="Black colonies" ~ "Black colony",
   #          `COLONY MORPHOLOGY ON XLD`=="Blackish colonies" ~ "Black colony",
   #          TRUE ~ `COLONY MORPHOLOGY ON XLD`)
  #) 
#NB: $CITRATE  "+" "-" "d" ->  intermediary : its a standard code : we keep it like that 

## Renaming the data entries in Isolate column
Cleaned_Labdata_Original <- Cleaned_Labdata_Original %>%
  mutate(
    Isolate = case_when(
      Isolate == "S.paratyphi A (Except Indole)" ~ "S.paratyphi A",
      Isolate == "S.typhi (Except Gas)" ~ "S.typhi",
      Isolate == "S.typhimariam" ~ "S.typhimurium",
      Isolate == "Shigella spp ( Except Gas)" ~ "Shigella spp",
      # Default: Keep all other values in the 'Isolate' column as they are
      TRUE ~ Isolate 
    )
  )

#Note I have silent the output here to not get enourmous amount of information in the output
#Cleaned_Labdata_Original %>%
#str(.)

```

```{r}
## Replacing NA in Isolate column with No growth
Cleaned_Labdata_Original$Isolate[is.na(Cleaned_Labdata_Original$Isolate)] <- "No growth"
#Check for duplicates
duplicated_test<-Cleaned_Labdata_Original$Isolate_ID[duplicated(Cleaned_Labdata_Original$Isolate_ID)]
print(duplicated_test)

# Get duplicated IDs (including first occurrences)
duplicated_test<-Cleaned_Labdata_Original$Isolate_ID[duplicated(Cleaned_Labdata_Original$Isolate_ID, fromLast = TRUE)]


# Remove duplicates from the list itself
unique_duplicated_ids <- unique(duplicated_test)

# Collapse into a single text string
text_string <- paste(unique_duplicated_ids, collapse = ", ")

# View result
print(text_string)

```

```{r}
CheckingDuplicates <- Cleaned_Labdata_Original%>%
  filter(Isolate_ID %in% 
           c("1392_1_D", "1395_1_D", "13150_2_D", "21303_2_R", "21345_1_R", "22287_1_R", "22315_1_R", "2374_1_D", "23130_1_D"))%>%
  select(INIKA_ID, Isolate_ID, Freezing_ID_No,Isolate,REGION, SEASON, `SAMPLE FROM` )
```

We detected that there are spelling mistakes in some of the Isolates_Ids as they are not corresponding to the correct INIKA_IDs. \# Solution: We make a new Isolate_ID calling it Isolate_ID_Corrected, but need to keep the wrong one as well in order to manage to join the data from TVLA (VITEK)

```{r}
# Get duplicated IDs (including first occurrences)
duplicated_test<-Cleaned_Labdata_Original$Isolate_ID[duplicated(Cleaned_Labdata_Original$Isolate_ID, fromLast = TRUE)]
```

```{r}
CheckingDuplicates <- Cleaned_Labdata_Original%>%
  filter(Isolate_ID %in% 
           c("1392_1_D", "1395_1_D", "13150_2_D", "21303_2_R", "21345_1_R", "22287_1_R", "22315_1_R", "2374_1_D", "23130_1_D"
          ))%>%
  select(INIKA_ID, Isolate_ID,Isolate,REGION, SEASON, `SAMPLE FROM`,`COLONY MORPHOLOGY ON BGA`, `COLONY MORPHOLOGY ON CARBA`, `COLONY MORPHOLOGY ON XLD`, `COLONY MORPHOLOGY ON C3GR` )
  
  # Remove duplicates from the list itself
unique_duplicated_ids <- unique(duplicated_test)
# Collapse into a single text string
text_string <- paste(unique_duplicated_ids, collapse = ", ")

# View result
print(text_string)
```

```{r}

Cleaned_Labdata_Original <- Cleaned_Labdata_Original %>%
  # 1. Group the data by Isolate_ID
  group_by(Isolate_ID) %>%
  # 2. Count the number of times each Isolate_ID appears
  mutate(Isolate_ID_repeat_count = n()) %>%
  # 3. Assign a sequential count (1, 2, 3...) within each Isolate_ID group
  mutate(Isolate_ID_row_count = row_number()) %>%
  # 4. Remove the grouping
  ungroup() %>%
  # 5. Modify the Isolate_ID based on the counts
  mutate(Isolate_ID = case_when(
    # The condition: If Isolate_ID appeared exactly 2 times AND it's the second row
    Isolate_ID_repeat_count == 2 & Isolate_ID_row_count == 2 ~
      # Action: Append the row count (2) to the Isolate_ID
      paste(Isolate_ID, Isolate_ID_row_count, sep = "_"),
    
    # The default: Otherwise, keep the original Isolate_ID
    TRUE ~ Isolate_ID
  ))


# We have now new Isolate_IDs as some were duplicates, but we will remove one further below.
Test<- Cleaned_Labdata_Original %>% 
  select(INIKA_ID, Isolate_ID,Isolate, Isolate_ID_repeat_count, Isolate_ID_row_count,
         REGION, SEASON, `SAMPLE FROM`,`COLONY MORPHOLOGY ON BGA`, `COLONY MORPHOLOGY ON CARBA`, `COLONY MORPHOLOGY ON XLD`, `COLONY MORPHOLOGY ON C3GR` ) %>% 
  filter(Isolate_ID_repeat_count=="2")
## Found one true duplicate, removed it below
```

Note! Madelaine changed below as the Freezing_ID_No=="2146-1_D" should be "2156_1_D", due to misspellings, not any from the SampleID called 2146 in the selected! We had to exclude the Isolate_ID 22315_1_R_2 as this is a true duplicate originating from the 22315, but mislabelleded in the original datafile to 22316_1_R instead of 22315_1_R, so when corrected in the original file we ending up with a true duplicate!

```{r}
Cleaned_Labdata_Original <- Cleaned_Labdata_Original%>% 
  mutate(Isolate_ID =case_when(Freezing_ID_No=="21161-1_D" ~ "21161_1_D",
                               Freezing_ID_No=="2146-1_D" ~ "2156_1_D",
                               TRUE ~ Isolate_ID,
                               ))%>% 
    filter(Isolate_ID!="22315_1_R_2")%>%
  filter(INIKA_ID !="21267")
  	


#Cleaned_Labdata_Original <- Cleaned_Labdata_Original %>%
 # mutate(
#    Isolate_ID_Corrected = case_when(
 #     Isolate == "C.freundii" & Isolate_ID == "2374_1_D" ~ "2374_2_D",
     # TRUE ~ Isolate_ID_Corrected
  #  )
 # )


```

From the next original file sheet 2 we get the Isolates that have been selected foir the VITEK confirmation at TVLA, Unfortunatel we should have selected also the CPE isolates which needs to be done separately, one of 11 has been selected through the process of systematic selection of every 5th E.coli.

Importing the DRY_RAINY_MWZ_KILIMANJARO_AST2 file, which includes the Selected isolates to be confirmed at TVLA by VITEK

```         
```

```{r}


# Use both parameters together
DRY_RAINY_MWZ_KILIMANJARO_AST2 <- read_excel(
  path = params$selection_file,
  sheet = params$selection_sheet
)

head(DRY_RAINY_MWZ_KILIMANJARO_AST2)


```

```{r}
# Selecting (Deselect) relevant columns
DRY_RAINY_MWZ_KILIMANJARO_AST2 <- 
  DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  select(-`S/No`, -`SAMPLE TYPE`,
         -`Season Included in Number? ( evt. how?)`,-Comment_)

library(dplyr)
library(stringr)

DRY_RAINY_MWZ_KILIMANJARO_AST2 <- DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  mutate(
    INIKA_ID = str_replace_all(INIKA_ID, "NIKA_", "INIKA_"),   # First replacement
    INIKA_ID = str_replace_all(INIKA_ID, "INIKA_", ""),        # Second replacement
    INIKA_ID = str_replace_all(INIKA_ID, "I", "")              # Third replacement
  ) %>%
  mutate_if(is.character, factor) %>%
  mutate(
    # Extract first part of INIKA_ID
    INIKA_prefix   = str_extract(INIKA_ID, "^[^_]+"),
    
    # Remove first part of Freezing_ID_No
    Isolate_suffix = str_replace(Freezing_ID_No, "^[^_]+_", ""),
    
    # Combine to make corrected ID
    Isolate_ID     = paste0(INIKA_prefix, "_", Isolate_suffix)
  )



#Check for duplicates
duplicated_test<-DRY_RAINY_MWZ_KILIMANJARO_AST2$Isolate_ID[duplicated(DRY_RAINY_MWZ_KILIMANJARO_AST2$Isolate_ID)]
print(duplicated_test)



```

```{r}
# 1. Group the data by Isolate_ID

DRY_RAINY_MWZ_KILIMANJARO_AST2 <- 
  DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
group_by(Isolate_ID) %>%
  # 2. Count the number of times each Isolate_ID appears
  mutate(Isolate_ID_repeat_count = n()) %>%
  # 3. Assign a sequential count (1, 2, 3...) within each Isolate_ID group
  mutate(Isolate_ID_row_count = row_number()) %>%
  # 4. Remove the grouping
  ungroup() %>%
  # 5. Modify the Isolate_ID based on the counts
  mutate(Isolate_ID = case_when(
    # The condition: If Isolate_ID appeared exactly 2 times AND it's the second row
    Isolate_ID_repeat_count == 2 & Isolate_ID_row_count == 2 ~
      # Action: Append the row count (2) to the Isolate_ID
      paste(Isolate_ID, Isolate_ID_row_count, sep = "_"),
    
    # The default: Otherwise, keep the original Isolate_ID
    TRUE ~ Isolate_ID
  ))


## Re labeling the data in the columns
# List the columns to be modified

cols_to_relabel <- c(
  "COLONY MORPHOLOGY ON C3GR", 
  "COLONY MORPHOLOGY ON CARBA", 
  "COLONY MORPHOLOGY ON XLD", 
  "COLONY MORPHOLOGY ON BGA"
) 

# Apply the replacement to the selected columns
DRY_RAINY_MWZ_KILIMANJARO_AST2 <- DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  mutate(
    across(
      .cols = all_of(cols_to_relabel),
      .fns = ~ replace(., . == "NBG", "No growth")
    )
  )

## Check the homogeneity of the data registering ----
# Trick to select all the columns of type character in DRY_RAINY_MWZ_KILIMANJARO_AST2 dataset
DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  select(where(is.character)) %>%
  # I modify all the columns from character to factor - because I want to see the "levels" : categories
  # to identify if you wrote homogeneously 
  mutate_all(factor) %>%
  # Select one column and look at the levels
  lapply(levels)

## Further cleaning for homogeneity entries
DRY_RAINY_MWZ_KILIMANJARO_AST2 <- 
  DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  # `SAMPLE TYPE` is only Stool BUT I do not want to remove because I want to combine with other dataset. 
  # `SAMPLE FROM` : Fix syntax  "Adult outPatient" "Adult outpatient"
  mutate(`SAMPLE FROM` = str_replace(`SAMPLE FROM`, "Adult outPatient", "Adult outpatient")) %>%
  # `NAME OF SCHOOL/HF` Homeogeneize :  "Pasua HC"  "PasuaHC"   From To 
  mutate(`NAME OF SCHOOL/HF` = str_replace(`NAME OF SCHOOL/HF`, "PasuaHC", "Pasua HC")) %>%
  #  `COLONY MORPHOLOGY ON C3GR` "Pinkish/Reddish" <- keep this one "pinkish/Reddish"
  mutate(`COLONY MORPHOLOGY ON C3GR` = str_replace(`COLONY MORPHOLOGY ON C3GR`, "pinkish/Reddish", "Pinkish/Reddish")) %>%
  # `COLONY MORPHOLOGY ON XLD`  "Black colonies"       "Black colony"         "Blackish colonies"  need to be identical
  ## "No black colony"      "No blackish colonies" must be identical - Fixed with several choices of equality
  mutate(`COLONY MORPHOLOGY ON XLD` = 
           case_when(
             `COLONY MORPHOLOGY ON XLD`=="No blackish colonies" ~ "No black colony",
             `COLONY MORPHOLOGY ON XLD`=="Black colonies" ~ "Black colony",
             `COLONY MORPHOLOGY ON XLD`=="Blackish colonies" ~ "Black colony",
             TRUE ~ `COLONY MORPHOLOGY ON XLD`)
  ) 
#NB: $CITRATE  "+" "-" "d" ->  intermediary : its a standard code : we keep it like that 

## Renaming the data entries in Isolate column
DRY_RAINY_MWZ_KILIMANJARO_AST2 <- DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  mutate(
    Isolate = case_when(
      Isolate == "S.typhimariam" ~ "S.typhimurium",
      # Default: Keep all other values in the 'Isolate' column as they are
      TRUE ~ Isolate 
    )
  )

```

```{r}
### Inserting new column "VITEK" ----
# Create the new column 'VITEK' and assign the value 1 to all rows, this means these have been selected for VITEK confirmation
Selection<- 
  DRY_RAINY_MWZ_KILIMANJARO_AST2 %>%
  mutate(VITEK = 1) %>%
  # Relocate the new column immediately after 'Isolate_ID'
  relocate(VITEK, .after = Isolate_ID)

```

Importing the original KOBOTOOLBOX data which contains the sampledata, demography

```{r}
library(readxl)
KOBOTOOLBOX_INIKA_SAMPLING_ANALYSIS_HUMAN_2025_06_12_18_24_20 <- read_excel("data/KOBOTOOLBOX_INIKA_SAMPLING_ANALYSIS_HUMAN_-_2025-06-12-18-24-20.xlsx",
             na = c("","NA"))


```

```{r}
#'Origin of sample' == "Outpatient"
#'Age (yrs)' == "45")
Check_KOBO<-KOBOTOOLBOX_INIKA_SAMPLING_ANALYSIS_HUMAN_2025_06_12_18_24_20 %>%
  filter(Region=="Kilimanjaro", District=="Moshi Rural", Season=="Wet" )

## Selecting relevant columns (removing unwanted columns) ----
Demographic_cleaned_data <-
  KOBOTOOLBOX_INIKA_SAMPLING_ANALYSIS_HUMAN_2025_06_12_18_24_20 %>%
  select(INIKA_OH_TZ_ID,`Age (yrs)`, Season, GPS_coordinates, `_GPS_coordinates_latitude`,  `_GPS_coordinates_longitude`,
         `_GPS_coordinates_precision`,Region, District, `Origin of sample`, `If school-aged children, name of the school*`,
         `Which class/grade are you?*`,`Who is your caretaker?*`,`If others, mention`,`What is your occupation and/or of your caretaker?*`,
         `Have you ever heard about AMR?`, `If yes, how did you get this information?`, `Have you or your children used any antibiotics at any time?`,
         `Have you or your children used any antibiotics in the past six months?`, `If yes, where did you get these drugs from?`,
         `If it was drug sellers or pharmacy; did you have a prescription from the doctor/prescriber?`,
         `If was prescribed by the prescriber, was it after Laboratory results of culture and sensitivity?`,
         `Do you have farm animals at your home place?`,`If Yes, do you participate in taking care of animals i.e. cleaning, Feeding?`, 
         `Are your animals being treated for any diseases?`, `Do you have a poultry farm at your home?`,
         `If yes, are they for eggs or meat?`,`Do you participate in taking care of Poultry i.e. Cleaning/Feeding?`,
         `Are your Poultry being treated by veterinary doctor?`, `If Yes, does the veterinary doctor tell you about the withdrawal time?`,
         `Do you have a toilet at your home place?`, `Do you wash your hands with soap after every toilet visit?`,Gender,`Origin of sample`,
         - GPS_coordinates) %>%
  rename("Latitude" = `_GPS_coordinates_latitude`,
         "Longitude" = `_GPS_coordinates_longitude`,
         "GPS_Precision" = `_GPS_coordinates_precision`,
         INIKA_ID = INIKA_OH_TZ_ID,
         Age_yrs = `Age (yrs)`,
         GENDER = `Gender`,
         REGION = `Region`,
         DISTRICT = `District`,
         SEASON = `Season`,
         ORIGIN_OF_SAMPLE = `Origin of sample`
         ) %>% 
  # Consider keeping as double
  mutate_at(.vars = "Age_yrs", as.character)


# Reordering the columns
Demographic_cleaned_data <- 
  Demographic_cleaned_data %>%
  select(INIKA_ID, 
         Age_yrs,
         GENDER, 
         SEASON,,
         REGION, DISTRICT,
         ORIGIN_OF_SAMPLE,
         Latitude, Longitude, GPS_Precision,
         everything())



# Get a logical vector: TRUE for elements that are duplicates (2nd, 3rd, etc. occurrences)
duplicates_logical <- duplicated(Demographic_cleaned_data$INIKA_ID)
#sum(duplicates_logical)

# Extract the actual duplicated ID values
duplicate_INIKA_IDs <- 
  Demographic_cleaned_data$INIKA_ID[duplicates_logical]

duplicate_INIKA_IDs
# Alternative - not better though
# Demographic_cleaned_data$INIKA_ID[
#   which(duplicated(Demographic_cleaned_data$INIKA_ID))
# ]


# View the unique duplicate ID values
unique_duplicate_INIKA_IDs <- unique(duplicate_INIKA_IDs)

# Print the result
print(unique_duplicate_INIKA_IDs)

#First check if the duplicated ID has exactly the same data
Demographic_cleaned_data %>%
  filter(INIKA_ID %in% unique_duplicate_INIKA_IDs) %>%
  arrange(INIKA_ID) %>%
  View()
```

Here your duplicates INIKA IDs contain different information : How can we know what is correct or not or what should we do ? keep in mind iff this one was sent to sequencing ?

HERE FIX or make an explicit choice ---- The INIKA_ID 21267 needs to be totally excluded as it is not possibe to know which of the persons that were sampled.

```{r}
#Remove the duplicates and maintain the fist occurrence
Demographic_cleaned_data <- 
  Demographic_cleaned_data %>%
  filter(INIKA_ID !=21267)%>%
  # Select only unique rows based on the INIKA_ID column
  # .keep_all = TRUE ensures that all other columns from the first 
  # row for each unique INIKA_ID are retained.
  distinct(INIKA_ID, .keep_all = TRUE)

Check_KOBO<-Demographic_cleaned_data %>%
  filter(REGION=="Kilimanjaro", DISTRICT=="Moshi Rural", SEASON=="Wet" )


# Madelaine:_transform INIKA_ID to charachter
Demographic_cleaned_data$INIKA_ID <- as.character(Demographic_cleaned_data$INIKA_ID)

# View a preview of the resulting data frame
head(Demographic_cleaned_data)

```

2204 unique Ids remained after removing one duplicate (21267)

Draw a table

Define Age Bins and Create the Summary Table

```{r}
summary_table <- Demographic_cleaned_data %>%
  mutate(Age_yrs=as.numeric(Age_yrs))%>%
   
   # Create the Age Range column using the 'Age_yrs' column
   mutate(`Age Range` = cut(Age_yrs,
                            breaks = c(10, 18, 28, 38, 48, 58, 68, 78, 89), # 89 ensures 78-88 is captured
                            labels = c('10 - 17', '18 - 27', '28 - 37', '38 - 47', 
                                       '48 - 57', '58 - 67', '68 - 77', '78 - 88'),
                            right = FALSE)) %>% # 'right = FALSE' makes the lower bound inclusive
   
   # Group by all requested demographic variables and the new age range
   group_by(REGION, DISTRICT, GENDER, SEASON, ORIGIN_OF_SAMPLE, `Age Range`) %>%
   
   # Count the number of observations (rows) for each group,
   #    as an explicit 'Total' column was not provided.
   summarise(`Total Count` = n()) %>%
   
   # Remove grouping structure
   ungroup()
 
 # Print the resulting table
 print(summary_table)
```

Optional: Write the resulting table to a new CSV file

```{r}
 
write_csv(summary_table, 
          "Results/Demographic_Summary_Table_Final_Test.csv")
```

```{r}
# Compute the Interquartile Range (IQR) 
# The na.rm = TRUE argument ensures that missing values are ignored in the calculation.
iqr_age <- IQR(Demographic_cleaned_data$Age_yrs, na.rm = TRUE)

# Print the result
print(paste("The Interquartile Range (IQR) of Age_yrs is:", iqr_age))

#library(dplyr)
#library(tidyr)
library(leaflet)

#Drawing a map
map_data <- Demographic_cleaned_data %>%
  # Rename columns 7 and 8 for clarity in the map code
  rename(
    latitude = 7,  # Column number 7 as Latitude
    longitude = 8  # Column number 8 as Longitude
  ) %>%
  # Ensure both columns are treated as numeric
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) %>%
  # Remove any rows where conversion failed or coordinates are missing (NA)
  filter(!is.na(latitude) & !is.na(longitude))

# 2. Draw the Interactive Map using Leaflet
map_leaflet <- map_data %>%
  leaflet() %>%
  # Add a base map layer (OpenStreetMap is a good default)
  addTiles() %>%
  # Add markers for each GPS coordinate in your dataset
  addMarkers(
    lat = ~latitude,
    lng = ~longitude,
    # Optional: Add a popup showing the coordinates when a marker is clicked
    popup = ~paste("Lat:", latitude, "<br>Lng:", longitude)
  )

# 3. Display the map
map_leaflet


```

```{r}
## Importing the original lab MALDITOF_RESULTS (From TVLA)
MALDITOF_RESULTS <- read_excel("data/MALDITOF_RESULTS.xlsx")
# Data cleaning ----
#Looking at the data structure ----
glimpse(MALDITOF_RESULTS)
MALDITOF_RESULTS[, "Isolate ID"]
str(MALDITOF_RESULTS)
names(MALDITOF_RESULTS)
```

```{r}
#Selecting relevant columns (removing unwanted columns) ----
MALDITOF_RESULTS_cleaned_data <-
  MALDITOF_RESULTS %>%
  select( "Isolate ID" , `Biochemica Id by Maulid`,"VITEK MS  Results")

# Renaming the column and name it TVLA_ID
MALDITOF_RESULTS_cleaned_data <-
  MALDITOF_RESULTS_cleaned_data %>%
  rename(VITEK_MS_Results = `VITEK MS  Results`,
         TVLA_ID = `Isolate ID`)

# Relocate the TVLA_ID, not necessary as we only have one TVLA_ID now!
#MALDITOF_RESULTS_cleaned_data <- MALDITOF_RESULTS_cleaned_data %>%
#relocate(TVLA_ID, .after = TVLA_ID)

## Replacing "-" with "_" in TVLA_ID column
MALDITOF_RESULTS_cleaned_data$TVLA_ID <- gsub("-", "_", MALDITOF_RESULTS_cleaned_data$TVLA_ID)

```

Re writing the entries in TVLA_ID to match with that in Original Lab data, seems to have been some mispellings in the data from TVLA, I suggest to call this corrected_TVLA_ID instead to not loose TVLA IDs, MN found mistakes in the file that have been joined in ). Join, so need to correct below 2374_1_D, 11301_1_R and 13122_1_D

```         
Re writing the entries in TVLA_ID to match with that in Original Lab data, 
seems to have been some mispellings in the data from TVLA
```

```{r}

MALDITOF_RESULTS_cleaned_data <-
  MALDITOF_RESULTS_cleaned_data%>%
 # some Isolates removed due to duplications of ID 
filter(!(TVLA_ID == "23372_1_R" & `Biochemica Id by Maulid` == "Klebsiella pneumoniae"))%>%
filter(!(TVLA_ID == "23224_1_R" & VITEK_MS_Results  == "No Identification" ))%>%
# Correcting misspellings and misidentification as far as possible
  mutate(Isolate_ID = case_when(
    TVLA_ID == "1152_1_D G.N" ~ "1152_1_D",
    TVLA_ID== "11321_(i)_R" ~ "11321_1_R",
    TVLA_ID== "11214(i)_R"  ~ "11214_1_R",
    TVLA_ID== "11253_1_R"  ~ "11252_1_R",
    TVLA_ID== "11262(i)_R"  ~ "11262_1_R",
    TVLA_ID== "11348(i)_R"  ~ "11348_1_R",
    TVLA_ID== "122_1_D"  ~ "125_1_D",
    TVLA_ID== "1268_1_R"  ~ "1268_1_D",
    TVLA_ID== "12132(i)_D"  ~ "12132_1_D",
    TVLA_ID== "21337_(i)_R"  ~ "21337_1_R",
    TVLA_ID== "23124_1_R"  ~ "23124_1_D",
    TVLA_ID== "25256_2_R"  ~ "23256_2_R",
    TVLA_ID== "23316_1R"  ~ "23316_1_R",
    TVLA_ID== "23237_R_"  ~ "23237_1_R",
    TVLA_ID== "13235_2_R"  ~ "13235_1_R",
    TVLA_ID== "12310_2_R"  ~ "12310_1_R",
    TVLA_ID== "12285_1_R_"  ~ "12285_1_R",
    TVLA_ID== "13122_1_D"  ~ "13123_1_D", # MN corrected it has to be in this direction
    TVLA_ID== "2337_1_R"  ~ "23370_1_R",
    TVLA_ID== "2216_1_D"  ~ "22161_1_D",
    TVLA_ID== "2369_1_D"  ~ "2366_1_D",
    TVLA_ID== "23347_2_"  ~ "23347_2_R",
    TVLA_ID== "23220_1_C"  ~ "23220_1_R",
    TVLA_ID== "23282_1_"  ~ "23282_1_R",
    TVLA_ID== "23308_1_R"  ~ "23308_2_R",
    TVLA_ID== "23320_2_"  ~ "23320_2_R",
    TVLA_ID== "21220_(i)_R"  ~ "21220_1_R",
    TVLA_ID== "21209_(i)_R"  ~ "21209_1_R",
    TVLA_ID== "2111_(i)_D"  ~ "2111_1_D",
    TVLA_ID== "12195(II)_R"  ~ "12195_2_R",
    TVLA_ID== "12195(i)_R"  ~ "12195_1_R",
    TVLA_ID== "11353_(i)_R"  ~ "11353_1_R",
    TVLA_ID== "11350(ii)"  ~ "11350_2_R",
    TVLA_ID== "11359(i)_R"  ~ "11359_1_R",
    TVLA_ID== "11308(i)_R"  ~ "11308_1_R",
    TVLA_ID== "11191(ii)_R"  ~ "11191_2_R",
    TVLA_ID== "C.1197_1_DMV"  ~ "1197_1_D",
    TVLA_ID== "11154mv"  ~ "11154_1_D",
    TVLA_ID== "1164_1_DMV"  ~ "1164_1_D",
    TVLA_ID== "12125_1_D"  ~ "12125_2_D",
    TVLA_ID== "2291_1_D"  ~ "2291_2_D",
    TVLA_ID== "1166_1_R"  ~ "1166_1_D",
    TVLA_ID== "12125_1_D"  ~ "12125_2_D",
    TVLA_ID== "23142_1_R"  ~ "23142_1_D",
    TVLA_ID== "23131-1-D"  ~ "23131_1_D",
    TVLA_ID== "12311_1_R"  ~ "12311_2_R",
    TVLA_ID== "2234_2_R"  ~ "22347_2_R",
    TVLA_ID== "2236_2_R"  ~ "22362_2_R",
    TVLA_ID== "212348_1_R" ~ "22348_2_R",
    TVLA_ID == "23308_2" ~ "23308_2_R_2",
    TVLA_ID == "12203_2"  ~ "12203_2_R",
    TVLA_ID == "11301-1-R"  ~ "11301_2_R", # MN new correction
    TVLA_ID == "2374-1-D"  ~ "2374_1_D", # MN new correction
    TVLA_ID == "2220_1_R"  ~ "22204_1_R", # MMJ new correction 23/10/25 and below
    TVLA_ID == "232606_2_R"  ~ "23260_2_R",  
    TVLA_ID == "11301_1_R"  ~ "11301_2_R",
    TVLA_ID == "21161_1_D"  ~ "21161_1_D",
    TVLA_ID == "21185_1_R"  ~ "21185_2_R",
    TVLA_ID == "21294_1_R"  ~ "21294_2_R",
    TVLA_ID == "23286_2_R"  ~ "23286_1_R",
    TVLA_ID == "12300_1_R"  ~ "12300_2_R",
    TVLA_ID == "12324_11_R"  ~ "12324_1_R",
    TVLA_ID == "22285_1_D"  ~ "22285_1_R",
    #TVLA_ID == "22131_2_D"  ~ "22131_1_D", # MN Note why was this one corrected it is numbered 22131_2_D in the WHONETdata! # But if this is the case we should keep it as it is!# MMJ; There is typing error, This number two isolates were isolated, Isolate 1 was E.coli and was not selected for AST, Isolate 2 was K.pneumoniae which was selected
    TVLA_ID == "2223_2_R"  ~ "22231_2_R",
    TVLA_ID == "22145_1_D"  ~ "22145_2_D",
    TVLA_ID == "22845_1_R"  ~ "22345_1_R",
    TVLA_ID == "12336_1_R"  ~ "12336_2_R",
    TVLA_ID =="21335_1_R" ~ "21235_1_R", # MN attached some more, correct?
    
    TRUE ~ TVLA_ID  # This is the "else" condition: keep any other values as they are
  ))%>%
  # Remove first part of Isolate_ID to use as the INIKA_ID for possible joining by two columns
  mutate(
    # Matches the first underscore (_) followed by any character (.) 
    # zero or more times (*), and removes the entire match.
    INIKA_ID = str_remove(Isolate_ID, "_.*")
  )

  MALDITOF_RESULTS_cleaned_data<-unique(MALDITOF_RESULTS_cleaned_data)

# Joining and checking until we managed to identify almost all that were not joined, used the helpfile ( where results was imported from the files below) in excel for checking

 ## Importing the MALDITOF results from NVI ## 02.03.2026 MMJ
  # Note here you can not import the first 3 rows!
  NVI_MALDITOF_results <- read_excel("data/MALDITOF_NVI 25feb2026.xlsx", skip=3)
  str(NVI_MALDITOF_results)
  

 
NVI_MALDITOF_results <- NVI_MALDITOF_results %>%
    rename(Isolate='Maldi-Tof. 25/2-26 tmf',Isolate_ID = Isolatnr) %>%
   mutate(`Isolate_ID` = str_replace_all(`Isolate_ID`, "-", "_"))%>%
  mutate(INIKA_ID=str_extract(`Isolate_ID`, "^[^_]+"))
         
# Combine the MALDITOF_RESULTS_cleaned_data and NVI_MALDITOF_results
  combined_MALDITOF_Results <- full_join(MALDITOF_RESULTS_cleaned_data, NVI_MALDITOF_results, by = c("Isolate_ID"))
 
  
   JoinedMALDITOF_Selected<-left_join(combined_MALDITOF_Results,DRY_RAINY_MWZ_KILIMANJARO_AST2,by = "Isolate_ID")
  notjoinedMALDITOF<-anti_join(combined_MALDITOF_Results,DRY_RAINY_MWZ_KILIMANJARO_AST2,by = "Isolate_ID")

  # We need to ensure we have one column that contain the INIKA_ID, now there is one #INIKA_ID.x, INIKA_ID.y and one INIKA_ID. if the INIKA_ID is missing we can use the #INIKA_ID.x and then we only keep the INIKA_ID.
 
    
JoinedMALDITOF_Selected <- JoinedMALDITOF_Selected %>%
  mutate(INIKA_ID = coalesce(INIKA_ID, INIKA_ID.x))%>%
  select (-c(INIKA_ID.x, INIKA_ID.y))


DuplicationsTVLA<-combined_MALDITOF_Results$Isolate_ID[duplicated(combined_MALDITOF_Results$Isolate_ID)] 
Duplications<-DuplicationsTVLA
#view(Duplications)

Selection<-DRY_RAINY_MWZ_KILIMANJARO_AST2
#JoinedSelected_TVLA<-left_join(DRY_RAINY_MWZ_KILIMANJARO_AST2,MALDITOF_RESULTS_cleaned_data,by = "Isolate_ID")
#FullJoinedSelected_TVLA<-full_join(DRY_RAINY_MWZ_KILIMANJARO_AST2,MALDITOF_RESULTS_cleaned_data,by = "Isolate_ID")%>%
#  select(INIKA_ID.x,TVLA_ID, Isolate_ID, Freezing_ID_No, 'Biochemica Id by Maulid',Isolate,VITEK_MS_Results )
#notjoinedSelected<-anti_join(DRY_RAINY_MWZ_KILIMANJARO_AST2,MALDITOF_RESULTS_cleaned_data,by = "Isolate_ID")
# 5 not joined

#VITEK<-MALDITOF_RESULTS_cleaned_data
VITEK<-JoinedMALDITOF_Selected


```

```         
```

```{r}
#Duplicates checking 

MALDITOF_RESULTS_duplicates <- MALDITOF_RESULTS_cleaned_data$Isolate_ID[duplicated(MALDITOF_RESULTS_cleaned_data$Isolate_ID)]
print("Actual duplicated Isolate_IDs (second or subsequent appearance):")
print(MALDITOF_RESULTS_duplicates) 

#JoinedTVLA_Selected_duplicates<-JoinedTVLA_Selected$Isolate_ID[duplicated
#(JoinedTVLA_Selected$Isolate_ID)]
#print("Actual duplicated Isolate_IDs (second or subsequent appearance):")

      
#Need to remove duplicates if they are true duplicates in MALDITOF_RESULTS_cleaned_data
MALDITOF_RESULTS_cleaned_data <- unique(JoinedMALDITOF_Selected)%>%
filter(TVLA_ID!="125_1_D")

```

Below we created new IDs for the duplicates so it will be possible to distinguish them, need to be careful about how these can be joined! as there will not be another with exactly the same ID in any of the original files, needs to be assigned in those files!

```{r}

```

```{r}
MALDITOF_RESULTS_cleaned_data<- MALDITOF_RESULTS_cleaned_data %>%
  # 1. Group the data by Isolate_ID
  group_by(Isolate_ID) %>%
  # 2. Count the number of times each Isolate_ID appears
  mutate(Isolate_ID_repeat_count = n()) %>%
  # 3. Assign a sequential count (1, 2, 3...) within each TVLA_ID group
  mutate(Isolate_ID_row_count = row_number()) %>%
  # 4. Remove the grouping
  ungroup() %>%
  # 5. Modify the Isolate_ID based on the counts
  mutate(Isolate_ID = case_when(
    # The condition: If Isolate_ID appeared exactly 2 times AND it's the second row
    Isolate_ID_repeat_count == 2 & VITEK_MS_Results != "Escherichia coli" ~
      # Action: Append the row count (2) to the Isolate_ID
      paste(Isolate_ID, Isolate_ID_row_count, sep = "_"),
    
    # The default: Otherwise, keep the original TVLA_ID
    TRUE ~ Isolate_ID
  ))

# Checking for the isolates that remains in the dataset after removing the duplicates, we could have removed the whole observation, but some of these might hav ebeen selected for further confirmation at NVI
Test<-MALDITOF_RESULTS_cleaned_data %>%
  filter(Isolate_ID%in% c( "21228_1_R" ,"23248_2_R" ,"23277_2_R", "23308_2_R", "13235_1_R" ,"23372_1_R","23224_1_R","23370_1_R"  ))

```

Note for one Klebsiella pneumoniae the TVLA_ID = 23308_1_R and the created Isolate_ID = 23308_2_R Note for one Escherichia coli the TVLA_ID = 13235_2_R and the created Isolate_ID= 13235_1_R Need to check that we manage to join these two isolates properly! ( Evt change the Isolate_ID where we know it is wrongly spelled!)

```{r}
## Creating new column named ESC indicating the isolates for AST
MALDITOF_RESULTS_cleaned_data <- MALDITOF_RESULTS_cleaned_data %>%
  mutate(
    ESC = if_else(
      VITEK_MS_Results %in% c("Escherichia coli", "Klebsiella pneumoniae"),
      1,
      0
    )
  )


# 3o/10/2025 Corrected one isolate_ID above 
# Selecting confirmed Isolate as ESC for the Cleaneddata where we have the AST data

AST_ISOLATES <- MALDITOF_RESULTS_cleaned_data %>% 
  filter(ESC == 1)
```

```{r}

## Duplicates checking
AST_ISOLATES_duplicates <- AST_ISOLATES$Isolate_ID[duplicated(AST_ISOLATES$Isolate_ID)]
print("Actual duplicated TVLA_IDs (second or subsequent appearance):")
print(AST_ISOLATES_duplicates) ## "23277_2_R"


```

Cleaning of the 4. script

```{r}
# Importing the original WHONET DATA
WHONET_DATA_9_9_2025 <- read_excel("data/WHONET DATA_9_9_2025.xlsx")

## Looking at the data structure ----
glimpse(WHONET_DATA_9_9_2025)
str(WHONET_DATA_9_9_2025)
names(WHONET_DATA_9_9_2025)
```

Selecting the relevant columns

```{r}

WHONET_Cleaned_data <-WHONET_DATA_9_9_2025 %>%
  select(-ROW_IDX, -LABORATORY, -COUNTRY_A, -PATIENT_ID, -X_SPECIES, -INSTITUT, -SPEC_TYPE, - X_AST_DATE, -SPEC_CODE, -X_SPEC_HUM, -X_SPEC_ANI, 
         -X_SPEC_ENV, -ISOL_NUM, -ORG_TYPE, -COMMENT, -DATE_DATA, -TYL_ED30, -ORIGIN)%>%

#Changing the numeric into character across the columns

#Use mutate and across to apply the change to multiple columns
  mutate(
    #Select all columns that are currently numeric 
    across(where(is.numeric), ~ as.character(.))
  )%>%

#Renaming the columns names
rename(Isolate_ID = `SPEC_NUM`,
       INIKA_ID = `X_INIKA_ID`,
       REGION = `X_REGION`,
       DISTRICT = `X_DISTRICT`,
       SEASON = `X_SEASON`,
       ORIGIN_OF_SAMPLE = `X_ORIGIN_T`,
       PROTOCOL = `X_PROTOCOL`)
 

```

We use the one above even if there are duplication on INIKA_ID to be joined, in script 9.Join

```{r}
# Duplicates checking in INIKA_ID column
WHONET_duplicates <- WHONET_Cleaned_data$INIKA_ID[duplicated(WHONET_Cleaned_data$INIKA_ID)]
print("Actual duplicated Isolate_IDs (second or subsequent appearance):")
print(WHONET_duplicates)
```

From the duplicates, "23346" "23272" have the same isolate "S.typhimurium", while the rest duplicated with different Isolates (species) "23272" have confirmed as E.coli by VITEK_MS

```{r}
#Duplicates checking in Isolate_ID column
WHONET_duplicates <- WHONET_Cleaned_data$Isolate_ID[duplicated(WHONET_Cleaned_data$Isolate_ID)]
print("Actual duplicated Isolate_IDs (second or subsequent appearance):")
print(WHONET_duplicates)
```

Under here done on 18/10/2025, by Maulid Inserting a column "ESBL_Selection" its contents being the difference between CTC_ED30 and CTX_ED5, If the results will be \>= 5 (ESBL) MN Comment, here there is a misunderstanding mm zones should be below 22mm\_ therefore Madelaine has modified the following:

```{r}
WHONET_Cleaned_data <- WHONET_Cleaned_data %>%
   mutate(
  CTC_ED30 = as.numeric(as.character(CTC_ED30)),
  CTX_ED5 = as.numeric(as.character(CTX_ED5)),
  CRO_ED5 = as.numeric(as.character(CRO_ED30)),

  
   ESBL_Selection = CTC_ED30 - CTX_ED5
   )
```

Select the true ESBL\_ \# MN This is to early to do here! You first need to select those that have mm zones for Cefotaxime below 22mm for E.coli, and 21 mm for Klebsiella pneumoniae Madelaine corrected 26.11 below! Note I have new names on the variables so we need to change throughout the scripts to ensure it works Note, I have renamed this outputfile to remain the same name! You don't need a new file here!

```{r}
WHONET_Cleaned_data <- WHONET_Cleaned_data %>%
  mutate(
    ESCR_ECO_presumptive = case_when(
      ORGANISM == "eco" & CTX_ED5  < 22 ~ 1,
      ORGANISM == "eco" & CRO_ED30 < 23  ~ 1,
      TRUE ~ 0
    ),
    ESCR_KPN_presumptive = case_when(
      ORGANISM == "kpn" & CTX_ED5 < 21 ~ 1,
      ORGANISM == "kpn" & CRO_ED30 < 23  ~ 1,
      TRUE ~ 0
    ),
    ESBL_Presumptivefinal = case_when(
      (ESCR_KPN_presumptive == 1 | ESCR_ECO_presumptive == 1) & ESBL_Selection >= 5 ~ 1,
      TRUE ~ 0
    )
  )


```

In total 190 ESCR Eco ( Presumptive), 170 ESBL Eco ( Presumptive), 148 ESCR K.Pneumoniae (Presumptive), 99 ESBL K.Pneumoinae ( Presumoptive), Note this one WHONET_Cleaned_data_ESBL has not been printed out!- you need to do that! For now I don't print out any files, we do that when we know all is fine!

```{r}
# Exporting the file as rds
#saveRDS(WHONET_Cleaned_data,"data/CLEANED_DATA/4_WHONET_Cleaned_data-2025-10-08.rds")
saveRDS(WHONET_Cleaned_data,"data/CLEANED_DATA/4_WHONET_Cleaned_data-2026-03-11.rds")

#write.csv(WHONET_Cleaned_data_ESBL,"data/CLEANED_DATA/WHONET_Cleaned_data_ESBL.csv")

# Below we don't do this any more it is included in the one above
#write.csv(WHONET_Cleaned_data_ESBL,"data/CLEANED_DATA/WHONET_Cleaned_data_ESBL_260311.csv")

```

Now all original files have been cleaned, and if anything occurs to not be joined in the last code for joining to a "final" dataset which should contain all the datasets together- we need to change accordingly in this script.

Dataset from Kobotoolbox_all samples with demographic data

```{r}
Kobo<-Demographic_cleaned_data
Kobo$INIKA_ID <- as.character(Kobo$INIKA_ID)

```

Dataset from labdata

```{r}
Labdata <- Cleaned_Labdata_Original
#Change to character from the INIKA_ID to Numeric in Labdata for compatible joining with Kobo data. By MMJ 21-10_2025
Labdata$INIKA_ID <- as.character(Labdata$INIKA_ID)
```

```{r}
#join Kobo Labdata

joined_data <- Kobo %>%
  left_join(Labdata, by = "INIKA_ID")

# Checked all have been joined
Checknotjoined<-anti_join(Labdata,Kobo, by="INIKA_ID")
```

Next file to join:Selection file

```{r}
# Join the joined_data with the Selection by using Isolate_ID
joined_data <- joined_data%>%
  left_join(Selection, by = "Isolate_ID")
# Checked all have been joined
Checknotjoined<-anti_join(Selection,joined_data, by="Isolate_ID")
```

```{r}
# Join the joined_data with WHONETdata by using Isolate_ID

joined_data <- joined_data %>%
  left_join(WHONET_Cleaned_data, by = "Isolate_ID")

Checknotjoined<-anti_join(WHONET_Cleaned_data,joined_data, by="Isolate_ID")
```

Until here it joins fine, problem with the next joining

```{r}
# Join joined_data with MALDITOF_RESULTS_cleaned_data by Isolate_ID
joined_data <- joined_data %>%
left_join(MALDITOF_RESULTS_cleaned_data, by = "Isolate_ID")%>% 
  distinct()
Checknotjoined<-anti_join(joined_data,MALDITOF_RESULTS_cleaned_data, by = "Isolate_ID")
```

Checking the rows where matching has occurred several times Check1\<-joined_data[193,] Check2\<-joined_data[146,]

25.10.25 I think we have a finaldataset to work further on from here! We could filter ( content from different columns ) and select those variables needed for the specific analyses, but also to have an additional check that all is ok!

joined_data \<- joined_data%\>% mutate(Isolate_ID = case_when(Isolate_ID == "23308_1_R" \~ "23308_3_R", Isolate_ID == "23308_2_R" \~ "23308_1_R", TRUE \~ Isolate_ID)) Save the file

```{r}
write.csv(joined_data,"data/CLEANED_DATA/joined_data.csv")
```
