---
title: "JOINING_all_Datasets_Humans"
author: "Madelaine Norström"
date: "2025-12-02"
output: html_document
params: 
  script_name: "codes/JOINING_all_Datasets_Humans.Rmd"
  vitek_ast_data: "data/MALDITOF_RESULTS.xlsx"
  biochemical_data: "data/DRY_RAINY_MWZ_KILIMANJARO.xlsx"
  whonet: "data/WHONET DATA_9_9_2025.xlsx"
  selection_file: "data/DRY_RAINY_MWZ_KILIMANJARO.xlsx"
  selection_sheet: "DRY_RAINY_MWZ_KILIMANJARO_AST2"
  kobo: "data/KOBOTOOLBOX_INIKA_SAMPLING_ANALYSIS_HUMAN_-_2025-06-12-18-24-20.xlsx"
---

```{r}
params
```

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(tidyverse)
library(readxl)
library(kableExtra)
library(here)


# Anchor the project root
here::i_am(params$script_name)
```
```{r}
#| echo: true
#| message: false
#| warning: false

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```
Before running this script the Cleaning_all_original_datasets.Rmd needs to be run!

```{r}
Kobo<-Demographic_cleaned_data
Kobo$INIKA_ID <- as.character(Kobo$INIKA_ID)
```
Dataset from labdata

```{r}
Labdata <- Cleaned_Labdata_Original
```




# Import the Cleaned_Labdata_Original
Labdata <- readRDS("data/CLEANED_DATA/1_Cleaned_Labdata_Original-2025-10-07.rds")%>% 
  filter(!INIKA_ID == "21267")

#Labdata<-Cleaned_Labdata_Original%>%
 # filter(!INIKA_ID == "21267")
 
# Change to character from the INIKA_ID to Numeric in Labdata for compatible joining with Kobo data. By MMJ 21-10_2025
Labdata$INIKA_ID <- as.character(Labdata$INIKA_ID)

#join Kobo Labdata

joined_data <- Kobo %>%
  left_join(Labdata, by = "INIKA_ID")
# %>% 
#   mutate(Isolate_ID = case_when(Isolate_ID == "23308_1_R" ~ "23308_3_R",
#                                 Isolate_ID == "23308_2_R" ~ "23308_1_R", 
#                                 TRUE  ~  Isolate_ID)) 
#   
# Checked all have been joined
Checknotjoined<-anti_join(Labdata,Kobo, by="INIKA_ID")
# 0 observations

# Importing the selection file
Selection <- readRDS("data/CLEANED_DATA/Selection-2025-10-07.rds")

# Join the joined_data with the Selection by using Isolate_ID
joined_data <- joined_data%>%
  left_join(Selection, by = "Isolate_ID")
# Checked all have been joined
Checknotjoined<-anti_join(Selection,joined_data, by="Isolate_ID")
# 0 observations

# joining with WHONETdata from 4. Cleaning_WHONET_Data
# Importing the WHONET_Cleaned_data
WHONETdata <- readRDS("data/CLEANED_DATA/4_WHONET_Cleaned_data-2025-10-08.rds")

# Join the joined_data with WHONETdata by using Isolate_ID

joined_data <- joined_data %>%
  left_join(WHONET_Cleaned_data, by = "Isolate_ID")
# NOte this is the previous code below as it wasdepending on importing the dataset 
#joined_data <- joined_data %>%
#  left_join(WHONETdata, by = "Isolate_ID")

Checknotjoined<-anti_join(WHONET_Cleaned_data,joined_data, by="Isolate_ID")


# Importing the MALDITOF_RESULTS_cleaned_data
MALDITOF_RESULTS_cleaned_data <- read.csv("data/CLEANED_DATA/MALDITOF_RESULTS_cleaned_data.csv")

# Join joined_data with MALDITOF_RESULTS_cleaned_data by Isolate_ID
joined_data <- joined_data %>%
left_join(MALDITOF_RESULTS_cleaned_data, by = "Isolate_ID")%>% 
  distinct()

## Select the true ESBL # MMJ added this 
joined_data <- joined_data %>%
  mutate(
    ESBL = if_else(ESBL_Selection >= 5, 1, 0, missing = 0)
  ) 

Checknotjoined<-anti_join(joined_data,MALDITOF_RESULTS_cleaned_data, by = "Isolate_ID")


#Checking the rows where matching has occurred several times
# Check1<-joined_data[193,] 
# Check2<-joined_data[146,] 

# 25.10.25 I think we have a finaldataset to work further on from here!
#We could filter ( content from different columns ) and select those variables needed for the specific analyses, but also to have an additional check that all is ok!

# joined_data <- joined_data%>% 
#   mutate(Isolate_ID = case_when(Isolate_ID == "23308_1_R" ~ "23308_3_R",
#                                 Isolate_ID == "23308_2_R" ~ "23308_1_R", 
#                                 TRUE  ~  Isolate_ID)) 
# Save the file
write.csv(joined_data,"data/CLEANED_DATA/joined_data.csv")

# Importing the joined_data
joined_data <- read.csv("data/CLEANED_DATA/joined_data.csv")
# Descriptive_LabStyphi<-joined_data%>%
#   filter(Isolate.x == "S.typhimurium")
# Descriptive_LabECO1<-joined_data%>%
#   filter(Isolate.x == "E.coli")
# Descriptive_LabKpn1<-joined_data%>%
#   filter(Isolate.x == "K.pneumoniae")

Confirmed_by_TVLA<-joined_data%>%
  filter(!is.na(TVLA_ID)) # 397 observations
###############################################################################
# we detected a duplication of Isolate_ID in the MALDITOF_RESULTS_cleaned_data and we need to remove 
# We also detected that the ISOLATE_ID that was 125_1_D was duplicated, however the TVLA had misspelled the one of the IDs 
# and we remove the one where it was not identification; NOTE In the datasets for comparison MN detected that 13122_1_D , 11301_1_R and 2374_1_D needs to be excluded! Wrong numbers!
#MALDITOF_RESULTS_cleaned_data<-unique(MALDITOF_RESULTS_cleaned_data)%>%
 # filter(!TVLA_ID %in% c("125-1-D", '23370-1-R','23308-2'),
      #   !(VITEK_MS_Results == "Enterobacter asburiae  and Enterobacter cloacae" & TVLA_ID == "21228-1-R"),
       #  !(VITEK_MS_Results == "Citrobacter Werkmanii" & TVLA_ID == "23248-2-R"),
        #  !(VITEK_MS_Results == "No Identification" & TVLA_ID == "23224-1-R"),
        # !(VITEK_MS_Results == "Citrobacter freundii" & TVLA_ID == "13235-1-R"),
        # !(VITEK_MS_Results == "Enterobacter cloacae and Enterobacter asburiae" & TVLA_ID == "23372-1-R"),
        #
         #  )


#Check<-MALDITOF_RESULTS_cleaned_data%>%
 # filter(TVLA_ID=="23372-1-R")

# Below we don't get the all the isolates to join, tried to use TVLA_ID for the 10 that did not join but then the INIKA_ID was not there for all
  
# joined_data <-left_join(MALDITOF_RESULTS_cleaned_data,joined_data, by = "Isolate_ID")
# NOT<- anti_join(MALDITOF_RESULTS_cleaned_data,joined_data, by = "Isolate_ID")
# 2395_2_D was not selected, has to be excluded, # 23308_2_R should be changed
#Joined_data_2<-left_join(NOT,joined_data, by = "TVLA_ID")

#NOT<- anti_join(NOT,Joined_data_2, by = "TVLA_ID")

names(joined_data)

# Not when not writing out I have not got the step where the names have been changed, so either do this here again or see where this is performed...

# "COLONY MORPHOLOGY ON C3GR.x"                                                                     
#[340] "COLONY MORPHOLOGY ON CARBA.x"                                                                    
#[431] "COLONY MORPHOLOGY ON XLD.x"                                                                      
#[42] "COLONY MORPHOLOGY ON BGA.x"                                                                      
#[43] "TSI Media_Slope.x"                                                                               
#[44] "TSI Media_Butt.x"                                                                                
#[45] "TSI Media_Gas.x"                                                                                 
#[46] "TSIMedia_H2S.x"                                                                                  
#[47] "SIM Media_H2S.x"                                                                                 
#[48] "SIM Media Indole.x"                                                                              
#[49] "SIM Media Motility.x"

Maulid_joined_almost_all<-joined_data%>%
  rename(COLONY.MORPHOLOGY.ON.C3GR =`COLONY MORPHOLOGY ON C3GR.x`)
select(INIKA_ID.x, Isolate_ID, TVLA_ID, Isolate.x, VITEK_MS_Results,REGION.x,
         DISTRICT.x,SEASON.x, ORIGIN_OF_SAMPLE.x,`COLONY.MORPHOLOGY.ON.C3GR.x`, `COLONY.MORPHOLOGY.ON.CARBA.x`, 
         `COLONY.MORPHOLOGY.ON.XLD.y`, `COLONY.MORPHOLOGY.ON.BGA.y`, CITRATE.y, `TSI.Media_Slope.x`, `TSI.Media_Butt.x`, 
         `TSI.Media_Gas.y`, TSIMedia_H2S.y, `SIM.Media_H2S.x`, `SIM.Media.Indole.x`,`SIM.Media.Motility.x`,
         UREASE.y)
names(Maulid_joined_almost_all)
# Save the file
write.csv(Maulid_joined_almost_all, "data/CLEANED_DATA/Maulid_joined_almost_all.csv")

# We need to see first the results for all the different Biochemical tests for the Isolate obtained originally
# Descriptive_Lab<-Maulid_joined_almost_all%>%
#   filter(Isolate.x == "S.typhimurium")
# Descriptive_LabECO<-Maulid_joined_almost_all%>%
#   filter(Isolate.x == "E.coli")
# Descriptive_LabKpn<-Maulid_joined_almost_all%>%
#   filter(Isolate.x == "K.pneumoniae")

long_data <- Maulid_joined_almost_all %>%
  pivot_longer(
    cols = c(
      `COLONY.MORPHOLOGY.ON.C3GR.y`,
      `COLONY.MORPHOLOGY.ON.CARBA.y`, 
      `COLONY.MORPHOLOGY.ON.XLD.y`, 
      `COLONY.MORPHOLOGY.ON.BGA.y`, 
      CITRATE.y, `TSI.Media_Slope.y`, 
      `TSI.Media_Butt.y`, 
      `TSI.Media_Gas.y`, TSIMedia_H2S.y, 
      `SIM.Media_H2S.y`, `SIM.Media.Indole.y`,
      `SIM.Media.Motility.y`,
      UREASE.y
    ),
    names_to = "Method",
    values_to = "Content"
  )

grouped_summary <- long_data %>%
  group_by(Isolate.x, VITEK_MS_Results.x, Method, Content) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  arrange(Isolate.x, Method)



# We filter out the columns where Biochemical and TVLA results are the same
ECO_correct_Identified<-grouped_summary%>%
  filter(Isolate.x =="E.coli", VITEK_MS_Results.x =="Escherichia coli") ## 130 observation

## Save the file_ Maulid needs toi complete. 
write_tsv(ECO_correct_Identified,"data/CLEANED_DATA/ECO_correct_Identified.tsv")

saveRDS(ECO_correct_Identified,"data/CLEANED_DATA/ECO_correct_Identified.rds")

ECO_NOT_correct_Identified_Bio<-grouped_summary%>%
  filter(Isolate.x =="E.coli", VITEK_MS_Results!="Escherichia coli") 

write_tsv(ECO_NOT_correct_Identified_Bio,"data/CLEANED_DATA/ECO_NOT_correct_Identified_Bio.tsv")

saveRDS(ECO_correct_Identified,"data/CLEANED_DATA/ECO_NOT_correct_Identified_Bio.rds")

ECO_NOT_correct_Identified_VITEK<-grouped_summary%>%
  filter(Isolate.x !="E.coli", VITEK_MS_Results =="Escherichia coli") 

write_tsv(ECO_NOT_correct_Identified_VITEK,"data/CLEANED_DATA/ECO_NOT_correct_Identified_VITEK.tsv")

saveRDS(ECO_NOT_correct_Identified_VITEK,"data/CLEANED_DATA/ECO_NOT_correct_Identified_VITEK.rds")
# X in agreement

Kleb_correct_Identified<-grouped_summary%>%
  filter(Isolate.x =="K.pneumoniae", VITEK_MS_Results =="Klebsiella pneumoniae") # 52 observation

## Save the file_ Maulid needs to complete. 
write_tsv(Kleb_correct_Identified,"data/CLEANED_DATA/Kleb_correct_Identified.tsv")

saveRDS(Kleb_correct_Identified,"data/CLEANED_DATA/Kleb_correct_Identified.rds") 

Kleb_NOT_correct_Identified_Bio<-grouped_summary%>%
  filter(Isolate.x =="K.pneumoniae", VITEK_MS_Results!="Klebsiella pneumoniae")

write_tsv(Kleb_NOT_correct_Identified_Bio,"data/CLEANED_DATA/Kleb_NOT_correct_Identified_Bio.tsv")

saveRDS(Kleb_NOT_correct_Identified_Bio,"data/CLEANED_DATA/Kleb_NOT_correct_Identified_Bio.rds")

Kleb_NOT_correct_Identified_VITEK<-grouped_summary%>%
  filter(Isolate.x !="K.pneumoniae", VITEK_MS_Results=="Klebsiella pneumoniae") # 13 observation

write_tsv(Kleb_NOT_correct_Identified_VITEK,"data/CLEANED_DATA/Kleb_NOT_correct_Identified_VITEK.tsv")

saveRDS(Kleb_NOT_correct_Identified_VITEK,"data/CLEANED_DATA/Kleb_NOT_correct_Identified_VITEK.rds")

#Write out all the tables correct and not correct


# From the total joined dataset _ we need to check how many Isolates you have per category


# Now we have to join with NVI_data as well

#Check<-joined_data[2503,] 
#Check<-joined_data[193,] 
################################################################################
